---
title: "Data"
author: "Brenden Bready"
date: "2/9/2022"
output:
  word_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE}
library(rvest)
library(tidyverse)
```


```{r}
url <- "https://www.transfermarkt.us/premier-league/gesamtspielplan/wettbewerb/GB1/saison_id/2021"
# file <- read_html(url)
```



```{r}
tables <- url %>% 
  read_html() %>%
  html_table(fill = TRUE)
tables <- tables[4:length(tables)]
```

```{r}
tableone <- tables[[1]]
```


```{r}
clean_tables <- function(first_table){
    names(first_table) <- c("date", "2", "home", "4", "result", "6", "away", "8", "9", "10", "11")
    first_table <- first_table %>%
        select(-c(2, 4, 6, 8, 9, 10, 11))
    first_table <- suppressWarnings(first_table %>%
                                        filter(., substr(result, 1, 1) == "-" | substr(home, 1, 1) == "(") %>%
                                        separate(., result, into = c("home_score", "away_score"), sep = ":") %>%
                                        separate(., date, into = c("time", "day"), sep = "\t\t\t\t\t\t\t\t\t\t", fill = "right") %>%
                                        select(-time) %>%
                                        fill(., day, dir("down")) %>%
                                        separate(., home, into = c("home_position", "home_team"), sep = "\\)") %>%
                                        separate(., away, into = c("away_team", "away_position"), sep = "\\("))
    if(is.na(first_table$home_team[1])){
        first_table <- first_table %>%
            rename(., home_team = home_position, home_position = home_team)
    }
    first_table <- first_table[, c("day", "home_team", "home_position", "home_score", "away_score", "away_team", "away_position")]
    first_table <- first_table %>%
        select(-c(home_position, away_position))
    first_table <- suppressWarnings(first_table %>%
                                        mutate(home_team = str_trim(home_team, "both"), away_team = str_trim(away_team, "both"), home_score = as.numeric(home_score), away_score = as.numeric(away_score), day = str_trim(day, "both")) %>%
                                        mutate(., day = as.Date(day, "%m/%d/%y")) %>%
                                        rename(., Date = day, 'Home team' = home_team, 'Away team' = away_team, 'Home score' = home_score, 'Away score' = away_score))
    return(first_table)
}
```

```{r}
finished_table <- function(tables){
  full_table <- data.frame(matrix(nrow = 0, ncol = 7))
  for (x in 1:length(tables)){
    temp_table <- tables[[x]]
    full_table <- rbind(full_table, clean_tables(temp_table))
  }
  return(full_table)
}
```

```{r}
full_table <- finished_table(tables)
```

```{r}
write.csv(full_table, file = "data.csv", row.names = FALSE)
```

Change blanks to NA in date column, then fill, then separate (possibly parse date?)

# Two piece error coming from date column (missing days)

```{r}
table_one <- clean_tables(tables[[1]])
```



think ab recreating standings in a table
track things like average goals scored, allowed, etc (put into standings table)
build function using complete data, then try to mid season
begin write up

to do:
fix errors
begin writing
standings table


```{r}
url2 <- "https://www.transfermarkt.us/wettbewerbe/europa"
file2 <- read_html(url2)
tables2 <- file2 %>% html_table(fill = TRUE)
leagues <- tables2[[1]]
```

```{r}
get_leagues <- function(leagues){
  leagues <- leagues %>%
    select(competition) %>%
    filter(row_number() %% 2 == 0) %>%
    mutate(., competition = tolower(competition)) %>%
    mutate(., competition = str_replace_all(competition, "[\\s\\.]", "-"))
  leagues$competition[23] = "nemzeti-bajnoksag"
  leagues$competition[8] = "super-lig"
country_code <- c("GB1", "ES1", "IT1", "L1", "FR1", "PO1", "RU1", "TR1", "NL1", "BE1", "UKR1", "A1", "GR1", "SC1", "C1", "KR1", "PL1", "TS1", "DK1", "SER1", "RO1", "SE1", "UNG1", "NO1", "ZYP1")
country <- c("England", "Spain", "Italy", "Germany", "France", "Portugal", "Russia", "Turkey", "Netherlands", "Belgium", "Ukraine", "Austria", "Greece", "Scotland", "Switzerland", "Croatia", "Poland", "Czech Republic", "Denmark", "Serbia", "Romania", "Sweden", "Hungary", "Norway", "Cyprus")
return(cbind(leagues, country, country_code))
}
write.csv(get_leagues(leagues), file = "country_options.csv", row.names = FALSE)
```

```{r}
test <- get_leagues(leagues)
```

```{r}
test$competition[which(test$country == "Germany")]
```




