---
title: "Modeling"
author: "Brenden Bready"
date: "3/31/2022"
output: html_document
---

```{r}
library(tidyverse)
```

```{r}
data <- read_csv("data.csv", show_col_types = FALSE)
```

```{r}
data <- data %>%
        rename(team = home_team, opponent = away_team, team_score = home_score, opp_score = away_score) %>%
        select(-c(home_position, away_position))
data2 <- data %>%
        rename(team = opponent, opponent = team, team_score = opp_score, opp_score = team_score)
double_table <- bind_rows(data, data2)
```

put in home indicator (first half 1, second half 0) in double table

for overall ratings, offensive / defensive

eventually get to win loss tie to get expected points

```{r}
double_table <- doubleTableFunc(data)
```


```{r}
mod1 <- glm(team_score ~ team + opponent + Location, data = double_table, family = poisson)
summary(mod1)
```

```{r}
exp(0.36-0.16+.25)
exp(0.36-0.38+.16)
```

```{r}
mod1$coefficients
```

```{r}
hockey <- read_csv("ECACresults.csv")
```

```{r}
hockeyMod <- glm(team_score ~ team + opponent + Location, data = hockey, family = poisson)
summary(hockeyMod)
```

```{r}
original <- data.frame(x = mod1$coefficients, rating = exp(mod1$coefficients))
original$team <- rownames(original)
row.names(original) <- NULL
original <- original %>%
  select(team, x, rating)
```


```{r}
defAVG <- (original %>%
  filter(str_detect(team, "opponent") == TRUE) %>%
  summarise(sum(x)/(n()+1)))[1,1]
offAVG <- (original %>%
  filter(str_detect(team, "team") == TRUE) %>%
  summarise(sum(x)/(n()+1)))[1,1]
intercept <- (original$x)[1]
```


```{r}
adjusted <- data.frame(team = unique(double_table$team))
adjusted <- data.frame(team = adjusted[order(adjusted$team),])
adjusted$Off = append(0, original$x[2:(length(original$x)/2)])
adjusted$Def = append(0, original$x[(length(original$x)/2+1):(length(original$x)-1)])
```

```{r}
adjusted$OffAdj = adjusted$Off+defAVG+intercept
adjusted$DefAdj = adjusted$Def+offAVG+intercept
adjusted$expOff = exp(adjusted$Off+defAVG+intercept)
adjusted$expDef = exp(adjusted$Def+offAVG+intercept)
```

```{r}
avgPoisOffCoef <- mean(adjusted$OffAdj)
avgPoisDefCoef <- mean(adjusted$DefAdj)
avgPoisCoef <- exp((avgPoisDefCoef+avgPoisOffCoef)/2)
```

```{r}
mean(adjusted$expOff)
mean(adjusted$expDef)
```



```{r}
scoreRate <- mean(double_table$team_score)
```




```{r}
# To get our C value
C = scoreRate/avgPoisCoef
```

```{r}
poissonOFFavg = mean(adjusted$expOff)
poissonDEFavg = mean(adjusted$expDef)
```


```{r}
# To get D value
D = sqrt(C*(poissonDEFavg/poissonOFFavg))
```


```{r}
adjusted$newOffExp = adjusted$expOff*D 
adjusted$newDefExp = adjusted$expDef*(C/D)
```

```{r}
mean(adjusted$newOffExp)
mean(adjusted$newDefExp)
```


```{r}
poissonFunc <- function(games){
  model <- glm(team_score ~ team + opponent + Location, data = games, family = poisson)
  original <- data.frame(x = model$coefficients, rating = exp(model$coefficients))
  original$team <- rownames(original)
  row.names(original) <- NULL
  original <- original %>%
    select(team, x, rating)
  defAVG <- (original %>%
    filter(str_detect(team, "opponent") == TRUE) %>%
    summarise(sum(x)/(n()+1)))[1,1]
  offAVG <- (original %>%
    filter(str_detect(team, "team") == TRUE) %>%
    summarise(sum(x)/(n()+1)))[1,1]
  intercept <- (original$x)[1]
  homeAdj <- original$rating[length(original$rating)]
  adjusted <- data.frame(team = unique(games$team))
  adjusted <- data.frame(team = adjusted[order(adjusted$team),])
  adjusted$Off = append(0, original$x[2:(length(original$x)/2)])
  adjusted$Def = append(0, original$x[(length(original$x)/2+1):(length(original$x)-1)])
  adjusted$OffAdj = adjusted$Off+defAVG+intercept
  adjusted$DefAdj = adjusted$Def+offAVG+intercept
  adjusted$expOff = exp(adjusted$Off+defAVG+intercept)
  adjusted$expDef = exp(adjusted$Def+offAVG+intercept)
  avgPoisOffCoef <- mean(adjusted$OffAdj)
  avgPoisDefCoef <- mean(adjusted$DefAdj)
  avgPoisCoef <- exp((avgPoisDefCoef+avgPoisOffCoef)/2)
  scoreRate <- mean(games$team_score)
  C = scoreRate/avgPoisCoef
  poissonOFFavg = mean(adjusted$expOff)
  poissonDEFavg = mean(adjusted$expDef)
  D = sqrt(C*(poissonDEFavg/poissonOFFavg))
  adjusted$newOffExp = adjusted$expOff*D 
  adjusted$newDefExp = adjusted$expDef*(C/D)
  returnedTable <- adjusted %>%
    mutate(OffRating = newOffExp, DefRating = newDefExp, OverallRating = newOffExp/newDefExp) %>%
    select(team, OffRating, DefRating, OverallRating) %>%
    arrange(desc(OffRating))
  returnedTable$OffRank = 1:nrow(returnedTable)
  returnedTable <- returnedTable %>%
    arrange(DefRating)
  returnedTable$DefRank = 1:nrow(returnedTable)
  returnedTable <- returnedTable %>%
    arrange(desc(OverallRating)) %>%
    mutate_if(is.numeric, round, digits=2) %>%
    select(team, OverallRating, OffRating, OffRank, DefRating, DefRank)
  return(list(returnedTable, homeAdj))
}
```


```{r}
poissonFunc(doubleTableFunc(data))[[2]]
```



```{r}
poissonFunc(hockey)
```



```{r}
doubleTableFunc <- function(games){
    data <- games %>%
        filter(!is.na(home_score & away_score))
    data <- data %>%
        rename(team = home_team, opponent = away_team, team_score = home_score, opp_score = away_score) %>%
        select(-c(home_position, away_position))
    data2 <- data %>%
        rename(team = opponent, opponent = team, team_score = opp_score, opp_score = team_score)
    double_table <- bind_rows(data, data2)
    double_table$Location = NA
    double_table$Location[0:length(double_table$date)/2] = 1
    double_table[is.na(double_table)] = -1
    return(double_table)
}
```

```{r}
dput(unique(poisson$team))
poisson <- poissonFunc(doubleTableFunc(data))
```

```{r}
poisson %>% 
  mutate(team = fct_reorder(team, (OverallRating))) %>%
  ggplot(aes(x = OverallRating, y = team)) + 
  geom_col(aes(fill = team)) +
  theme_bw() +
  xlab("Overall Rating") + 
  ylab("Team")
```

```{r}
rpi <- rpi((data))
rpi %>% 
  mutate(Team = fct_reorder(Team, (RPI))) %>%
  ggplot(aes(x = RPI, y = Team)) + 
  geom_col(aes(fill = Team)) +
  theme_bw(plot.title("RPI")) +
  xlab("RPI") + 
  ylab("Team")
```


Offensive Rating
Defensive Rating
Off / Def
Sort by overall rating
round decimals
add rank for offense and defense
record in standings function?
rating, offense, off rank, defense, def rank
draft of poster

```{r}
bb <- read_csv("BBstats.csv")
poissonFunc(doubleTableFunc(bb))
```

```{r}
rpi(bb)
```




international
football
ratings
prediction









