---
title: "RPI"
author: "Brenden Bready"
date: "3/8/2022"
output: html_document
---

```{r}
library(tidyverse)
```

```{r}
data <- read_csv("data.csv", show_col_types = FALSE)
```

```{r}
home_points <- data %>%
    group_by(home_team) %>%
    mutate(points = ifelse(home_score > away_score, 3, ifelse(home_score ==    away_score, 1, 0)))  %>%
    summarise(home_points = sum(points, na.rm=TRUE), home_games = sum(!is.na(points)))
away_points <- data %>%
    group_by(away_team) %>%
    mutate(points = ifelse(away_score > home_score, 3, ifelse(away_score == home_score, 1, 0))) %>%
    summarise(away_points = sum(points, na.rm=TRUE), away_games = sum(!is.na(points)))
teams_ppg <- cbind(home_points, away_points) %>%
  select(-c(away_team)) %>%
  rename(team = home_team) %>%
  mutate(ppg = (home_points+away_points)/(home_games+away_games))
```

```{r}
data <- data %>%
        rename(team = home_team, opponent = away_team, team_score = home_score, opp_score = away_score) %>%
        select(-c(home_position, away_position))
data2 <- data %>%
        rename(team = opponent, opponent = team, team_score = opp_score, opp_score = team_score)
double_table <- bind_rows(data, data2)
```


join above chunk to full games data file

ex for arsenal
have arsenal ppg
then get the average ppg for all teams arsenal plays from big dataset
then get average ppg for oponents of the teams that arsenal has played
assign weights to these values and average







Try a different approach here
Ignore and home and away and make a dataset double the length, just having a team column and opponent column
then do averages way easier


```{r}
ppg1 <- left_join(double_table, teams_ppg, by = "team") %>%
  select(-c(home_points, away_points, home_games, away_games)) %>%
  rename(team_ppg = ppg)
ppg2 <- left_join(double_table, teams_ppg, by = c("opponent" = "team")) %>%
  select(-c(home_points, away_points, home_games, away_games)) %>%
  rename(opp_ppg = ppg)
full_ppg <- left_join(ppg1, ppg2)
```

```{r}
opp_ppg <- full_ppg %>%
  group_by(team) %>%
  mutate(avg_opp_ppg = mean(opp_ppg)) %>%
  distinct(team, .keep_all = T) %>%
  select(c(team, avg_opp_ppg))
```

```{r}
new_teams_ppg <- left_join(teams_ppg, opp_ppg, by = c("team"))
```

```{r}
ppg4 <- left_join(double_table, new_teams_ppg, by = c("opponent" = "team")) %>%
  select(-c(home_points, away_points, home_games, away_games)) %>%
  rename(opp_ppg = ppg)
```

```{r}
new_opp_ppg <- ppg4 %>%
  group_by(team) %>%
  mutate(avg_opp_opp_ppg = mean(avg_opp_ppg)) %>%
  distinct(team, .keep_all = T) %>%
  select(c(team, avg_opp_opp_ppg))
```

```{r}
full_teams_ppg <- left_join(new_teams_ppg, new_opp_ppg, by = c("team")) %>%
  mutate(rpi = (ppg*(1/3)+avg_opp_ppg*(1/3)+avg_opp_opp_ppg*(1/3))) %>%
  arrange(desc(rpi))
```

full function

```{r}
rpi <- function(data, w1=(1/4), w2=(1/2), w3=(1/4)){
    data <- data %>%
        filter(!is.na(home_score & away_score))
    home_points <- data %>%
        group_by(home_team) %>%
        mutate(points = ifelse(home_score > away_score, 3, ifelse(home_score ==    away_score, 1, 0)))  %>%
        summarise(home_points = sum(points, na.rm=TRUE), home_games = sum(!is.na(points)))
    away_points <- data %>%
        group_by(away_team) %>%
        mutate(points = ifelse(away_score > home_score, 3, ifelse(away_score == home_score, 1, 0))) %>%
        summarise(away_points = sum(points, na.rm=TRUE), away_games = sum(!is.na(points)))
    teams_ppg <- cbind(home_points, away_points) %>%
        select(-c(away_team)) %>%
        rename(team = home_team) %>%
        mutate(ppg = (home_points+away_points)/(home_games+away_games))
    data <- data %>%
        rename(team = home_team, opponent = away_team, team_score = home_score, opp_score = away_score)
    data2 <- data %>%
        rename(team = opponent, opponent = team, team_score = opp_score, opp_score = team_score)
    double_table <- bind_rows(data, data2)
    ppg1 <- left_join(double_table, teams_ppg, by = c("opponent" = "team")) %>%
        select(-c(home_points, away_points, home_games, away_games)) %>%
        rename(opp_ppg = ppg)
    opp_ppg <- ppg1 %>%
        group_by(team) %>%
        mutate(avg_opp_ppg = mean(opp_ppg)) %>%
        distinct(team, .keep_all = T) %>%
        select(c(team, avg_opp_ppg))
    new_teams_ppg <- left_join(teams_ppg, opp_ppg, by = c("team"))
    ppg2 <- left_join(double_table, new_teams_ppg, by = c("opponent" = "team")) %>%
        select(-c(home_points, away_points, home_games, away_games)) %>%
        rename(opp_ppg = ppg)
    new_opp_ppg <- ppg2 %>%
        group_by(team) %>%
        mutate(avg_opp_opp_ppg = mean(avg_opp_ppg)) %>%
        distinct(team, .keep_all = T) %>%
        select(c(team, avg_opp_opp_ppg))
    full_teams_ppg <- left_join(new_teams_ppg, new_opp_ppg, by = c("team")) %>%
        mutate(rpi = (ppg*w1+avg_opp_ppg*w2+avg_opp_opp_ppg*w3)) %>%
        arrange(desc(rpi)) %>%
        select(c(team, rpi, ppg, avg_opp_ppg, avg_opp_opp_ppg, home_points, home_games, away_points, away_games)) %>%
        rename(Team = team, RPI = rpi, 'Points per game (PPG)' = ppg, 'Opponent\'s PPG' = avg_opp_ppg, 'Opponent\'s Opponent\'s PPG' = avg_opp_opp_ppg, 
               'Home points' = home_points, 'Home games' = home_games, 'Away points' = away_points, 'Away games' = away_games) %>%
        mutate_if(is.numeric, round, digits=2)
    return(full_teams_ppg)
}
```


```{r}
rpi(data)
```

Add function to shiny app in new tab
rearrange order

make double games into a function to help with poisson - do something

add poisson markdown with estimates and exponential

```{r}
mod1 <- glm(team_score ~ team + opponent, data = double_table, family = poisson)
summary(mod1)
```

```{r}
exp(0.36-0.16+.25)
exp(0.36-0.38+.16)
```









